TR Invariant T.I.

### 2017-04-21 Notes of various papers/TR invariant topological insulators T. Hughes. Dissertation/note.tex
[major] Various plots for infinite strip geometry added.

This amounts to removing the $A$ in top right corner and $C$ in bottom
left corner of eq.\ref{eq:disc-eigeneq}. The calculation can be found
in file \texttt{Lattice Dirac Model (2+1)-d-4-Infinite Stripe.nb}.
However, the solution by Mathematica is analytically intractable and
not really illuminating. Also plotting all those eigenvectors becomes
time-consuming when $N$ is large. So I looked for similar things
inside the book by Bernevig \cite{Bernevig2013}, and found the
following model.

@@

    \subsubsection{Edge States}
    \label{sec:Edge States}
    Inspired by lecture by Kane of Topological Insulators (BSS 2016), I
    will analyse the continuum limit of this model at the  point
    $(p_x=0,p_y=0)$, where the energy $E=0$. I also assume that we have placed
    to materials adjacent to each other in $x$ direction, on with
    $m<0$, and the other with $m>0$. Therefore, at the interface, we
    could have $m$ as a function of $x$ such that $m(0)=0$.

    At this point $x=0$, $\sin(p_x)\to p_x$ , and is replaced by $-i\hbar
    \partial _x$. So we recovered the continuum model:
    \begin{equation}
        H = -i\hbar\partial_x \sigma^x 
        -i\hbar\partial_y \sigma^y -m \sigma^z
    \end{equation}
    The Schrodinger equation is $H\psi =E\psi = 0$ for $E\approx 0$
    around this point. After some calculations, the equation is (with
    $\hbar=1$):
    \begin{equation}
        \left[i\partial_x \sigma^x + i\partial_y \sigma^y +
        m\sigma^z\right] \psi = 0
    \end{equation}
    However, this coupled PDE is hard to solve by Mathematica
    \footnote{See file \texttt{Lattice Dirac Model
    (2+1)-d-3-EdgeStates.nb} for my calculations.}. Therefore, I
    restrict their value on $x$, and solve the ODE:
    \begin{align}
        i\partial_x \psi_2(x) + m\psi_1(x) &= 0 \\
        i\partial_x \psi_1(x) - m\psi_2(x) &= 0
    \end{align}
    If assuming $m(x)$ is in the form of $m(x)=x$, i.e. positive when
    $x>0$, and negative when $x<0$, then the solution is:
    \begin{align}
        \psi_1 (x,y) &= e^{\text{Int}(x)} C(y) \\
        \psi_2 (x,y) &= -i \psi_1(x,y)
    \end{align}
    where $\text{Int}(x)=\int_1^x -m(k)\dd{k}$. $\text{Int}(x)$ has the
    property of goes to $-\infty$ as $x\to \pm\infty$.

@@

    It turns out that $(i,1)^t$ is also the eigenvector of
    $\sigma_y$ with eigenvalue $-1$. So the equation becomes quite
    simple now:
    \begin{equation}
        i\partial_y\chi(y) = E\chi(y)
    \end{equation}
    The solution is $\chi(y)=e^{-iEy}$, a running wave in the $y$
    direction!
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\bibliography{../../Library}{}
\bibliographystyle{alphaurl}

>>>>

@@

This amounts to removing the $A$ in top right corner and $C$ in bottom
left corner of eq.\ref{eq:disc-eigeneq}. 
\footnote{The calculation can be found
in file \texttt{Lattice Dirac Model (2+1)-d-4-Infinite Stripe
(N=3).nb}, and file \texttt{Lattice Dirac Model (2+1)-d-4-Infinite
Stripe (N=20).nb}.}
In this case, the crossing bands exists at slightly different value of
$m$, also, there are few sharp transitions from a closed gap state, to an
open gap state. For example, when $N=3$, the band crossing happens at
two range of $m$: from $0.65$ to $1.35$ (shown in
fig~\ref{fig:InfiniteStrip=3-1.jpg}, and from $2.4$ to $3.75$ (shown
in fig~\ref{fig:InfiniteStrip=3-2.jpg}). The cases when $N=10$,
$N=20$, $N=30$ also exhibits similar behavior. They also have almost
two ranges of values for $m$ when two bands close at one point. Their
result are plotted are at fig~\ref{fig:InfiniteStrip=10.jpg},
fig~\ref{fig:InfiniteStrip=20.jpg}, and ... the result for $N=30$ will
be updated here when my program finishes running.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=3-1}.jpg}
    \caption{Infinite Strip, $N=3$, closes at $p_x=0$}
    \label{fig:InfiniteStrip=3-1.jpg}
\end{figure}
\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=3-2}.jpg}
    \caption{Infinite Strip, $N=3$, closes at $p_x=\pi$}
    \label{fig:InfiniteStrip=3-2.jpg}
\end{figure}
\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=10}.jpg}
    \caption{Infinite Strip, $N=10$}
    \label{fig:InfiniteStrip=10.jpg}
\end{figure}
\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=20}.jpg}
    \caption{Infinite Strip, $N=20$}
    \label{fig:InfiniteStrip=20.jpg}
\end{figure}

As we have seen, the solution by Mathematica is analytically
intractable as $N$ grows larger. Also it would be too time-consuming
if I were to plotting those eigenvectors when $N$ is large. So I
looked for similar things inside the book by Bernevig
\cite{Bernevig2013}, and found the following model.

    \subsubsection{Edge States Analytically Obtained}
    \label{sec:Edge States}
    I will analyse the continuum limit of this model at the point
    $(p_x=0,p_y=0)$, where the energy $E=0$. I also assume that we
    have placed to materials adjacent to each other in $x$ direction.
    I will assume that the two materials are in different topological
    states. For example, if the band cross when $m=0$, then one
    material will have $m<1$, and the other will have $m>0$.
    Therefore, at the interface, we could have $m$ as a function of
    $x$ such that $m(0)=0$.

    This assumption implies that we break translational in both
    direction. So at this point $p_x=0$, $\sin(p_x)\to p_x$ , and is
    replaced by $-i\hbar \partial _x$. Similar for $\sin(p_y)$. So we
    recovered the continuum model:
    \begin{equation}
        H = -i\hbar\partial_x \sigma^x 
        -i\hbar\partial_y \sigma^y -m \sigma^z
    \end{equation}
    The Schrodinger equation is $H\psi =E\psi = 0$ for $E\approx 0$
    around this point. After some calculations, the equation is (with
    $\hbar=1$):
    \begin{equation}
        \left[i\partial_x \sigma^x + i\partial_y \sigma^y +
        m\sigma^z\right] \psi = 0
    \end{equation}
    However, this coupled PDE is hard to solve by Mathematica
    \footnote{See file \texttt{Lattice Dirac Model
    (2+1)-d-3-EdgeStates.nb} for my calculations.}. Therefore, I
    restrict their value on $x$, and solve the ODE:
    \begin{align}
        i\partial_x \psi_2(x) + m\psi_1(x) &= 0 \\
        i\partial_x \psi_1(x) - m\psi_2(x) &= 0
    \end{align}
    If assuming $m(x)$ is in the form of $m(x)=x$, i.e. positive when
    $x>0$, and negative when $x<0$, then the solution is:
    \begin{align}
        \psi_1 (x,y) &= e^{\text{Int}(x)} C(y) \\
        \psi_2 (x,y) &= -i \psi_1(x,y)
    \end{align}
    where $\text{Int}(x)=\int_1^x -m(k)\dd{k}$. $\text{Int}(x)$ has the
    property of goes to $-\infty$ as $x\to \pm\infty$.

@@

    It turns out that $(i,1)^t$ is also the eigenvector of
    $\sigma_y$ with eigenvalue $-1$. So the equation becomes quite
    simple now:
    \begin{equation}
        i\partial_y\chi(y) = E\chi(y)
    \end{equation}
    The solution is $\chi(y)=e^{-iEy}$, a free electron in the $y$
    direction!

    \textbf{Remark}: Although I have broken translational symmetry in
    both direction, I could do almost the same calculation with $p_y$
    conserved, e.g. in a infinite strip. In that case, I would get a
    state localized in the interface, and have $p_y=-E$, representing
    an electron travelling with momentum $-E$ in $y$'s direction.
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\bibliography{../../Library}{}
\bibliographystyle{alphaurl}
